\vspace{-1em}Guidance, please! Towards a framework for RDF-based constraint languages.

### 2015-07-16 paper/paper 2/dc-2015.tex
3.4 3.5

\section{Motivation}

Even with an upcoming W3C recommendation, it can be expected that several constraint languages will be used in practice in future -- consider the situation in the XML world, where a standardized schema language was available from the beginning and yet additional ways to formulate and check constraints have been created. Therefore, 
semantically equivalent \emph{specific constraints} represented in different languages will exist.
This raises two questions: 
\begin{enumerate}
 \item How can we ensure that two semantically equivalent constraints are actually validated consistently?
 \item How can we support the transformation of semantically equivalent constraints from one contraint language to another?
\end{enumerate}

@@

\subsection{\emph{Simple Constraints}}

@@

Constraints of 36\% of the constraint types are not expressible in DL but can still be described using the RDF-CV such as constraints of the type \emph{literal pattern matching} (\emph{R-44}) which restrict literals to match given patterns. The \emph{universal quantification (R-91)} {\small\ms{Book $\sqsubseteq$ $\forall$ identifier.ISBN}} ensures that books can only have valid \emph{ISBN} identifiers, i.e., strings that match a given regular expression.
%Even though, the restriction of the datatype \emph{ISBN} cannot be expressed in DL, \emph{OWL 2 DL} can be used to express the \emph{literal pattern matching} constraint:
%\begin{ex}
%ISBN a RDFS:Datatype ; owl:equivalentClass [ a RDFS:Datatype ;
    %owl:onDatatype xsd:string ; 
    %owl:withRestrictions ([ xsd:pattern "^\d{9}[\d|X]" ])] .
%\end{ex} 
%The first OWL 2 axiom explicitly declares {\em ISBN} to be a datatype. %\er{Why start from second and then first? Also better we say "owl axiom" when we talk axiom of owl, to prevent ambiguity.}\tb{resolved}
%The second OWL 2 axiom defines {\em ISBN} as an abbreviation for a datatype restriction on {\em xsd:string}. 
%The datatype {\em ISBN} can be used just like any other datatype like in the \emph{universal restriction} above.
%There are multiple use cases associated with the requirement to match literals according to given patterns (\emph{Literal Pattern Matching}\footnote{Corresponds to \emph{R-44-PATTERN-MATCHING-ON-RDF-LITERALS}}).
%The enterprise vessel, e.g.,  can only have the registry numbers "NCC-1701", "NCC-1701-A", "NCC-1701-B", "NCC-1701-C", "NCC-1701-D", or "NCC-1701-E".
%The universal restriction can be represented in DL:
%\ms{Enterprise $\sqsubseteq$ $\forall$ registryNumber.RegistryNumber}.
%The restriction of the datatype \ms{RegistryNumber}, however, cannot be expressed in DL, but OWL 2 DL can be used anyway to express the literal pattern matching constraint:
%
%\begin{ex}
%RegistryNumber
    %a rdfs:Datatype ; owl:equivalentClass [ a rdfs:Datatype ;
        %owl:onDatatype xsd:string ;
        %owl:withRestrictions ( [ xsd:pattern "NCC-1701([-][A-E])?" ] ) ] .
%\end{ex}
%The second axiom defines \ms{RegistryNumber} as an abbreviation for a datatype restriction on \ms{xsd:string}. 
%The first axiom explicitly declares \ms{RegistryNumber} to be a datatype. 
%The datatype \ms{RegistryNumber} can be used just like any other datatype like in the universal restriction above.
%The \emph{literal pattern matching} constraint validates \emph{ISBN} literals according to the regular expression causing constraint violations for triples which do not not match. 
%\ms{Janeway commandsEnterprise Voyager} and \ms{Voyager registryNumber "NCC-74656"\textasciicircum{}\textasciicircum{}RegistryNumber}, 
%but not for the triples \ms{Picard commandsEnterprise Enterprise} and \ms{Enterprise registryNumber "NCC-1701-E"\textasciicircum{}\textasciicircum{}RegistryNumber}.

Even though, constraints of the type \emph{literal pattern matching} cannot be expressed in DL, OWL 2 can be used to formulate this constraint:

@@

Table \ref{tab:simple-constraint-not-expressible-in-dl)} presents (1) how the not in DL expressible literal pattern matching constraint and (2) how the in DL expressible universal quantification are both represented using the RDF-CV. Thereby, the context class \emph{ISBN}, whose instances must satisfy the literal pattern matching constraint, is reused within the list of classes the universal quantification refers to. The literal pattern matching constraint type introduces the constraining element \emph{REGEX} whose validation has to be implemented once like for any other constraining element.
%In general, validation has to be implemented once for each \emph{generic constraint type} which is not expressible in DL. 
%\begin{gcotable}
%property & Enterprise & registryNumber & - & RegistryNumber & $\forall$ & - \\
%class & RegistryNumber & - & - & xsd:string & REGEX & 'NCC-1701([-][A-E])?' \\
%\end{gcotable}
\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Simple Constraints which are not Expressible in DL}
	\label{tab:simple-constraint-not-expressible-in-dl)}
	\centering
		\begin{tabular}{c|c|c|c|c|c}
      \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
      ISBN & - & - & xsd:string & REGEX & '\string^\text{$\backslash$d$\{9\}$[$\backslash$d$\mid$X]}\$' \\
      Book & identifier & - & ISBN & universal quantification & - \\
		\end{tabular}
\end{table} %\tb{Erman, can you please format the regular expression in the table above correctly?}

@@

\subsection{Bidirectional Constraint Transformations}

Constraint of the type \emph{minimum qualified cardinality restriction (R-75)} expressed in OWL 2:

@@

The same constraint expressed generically using the RDF-CV:

@@

SPIN inference rule to infer the generic constraint out of the constraint expressed by OWL 2:

@@

The property \emph{spin:rule} can be used to link an rdfs:Class with SPARQL CONSTRUCT queries. Each query defines an inference rule that is applied to all instances of the associated class and its subclasses. The inference rule defines how additional triples can be inferred from what is stated in the WHERE clause. For each binding of the pattern in the WHERE clause of the rule, the triple templates from the CONSTRUCT clause are instantiated and added as inferred triples to the underlying model. At query execution time, the SPARQL variable ?this is bound to the current instance of the class. As each resource per default is assigned to the class owl:Thing, this inference rule is evaluated for each subject of the input RDF graph.

SPIN inference rule to infer the OWL 2 constraint out of the generically expressed constraint:

\subsection{XXX}

@@

@@

The \emph{Dublin Core Application Profile (DCAP)} and \emph{Bibframe} are approaches to specify \emph{profiles} for application-specific purposes. The term \emph{profile} is widely used to refer to a document that describes how standards or specifications are deployed to support the requirements of a particular application, function, community, or context. In the metadata community, the term \emph{application profile} has been applied to describe the tailoring of standards for specific applications. A \emph{Dublin Core Application Profile (DCAP)} \cite{DCMI-DCAP-2009} defines metadata records which meet specific application needs while providing semantic interoperability with other applications on the basis of globally defined vocabularies and models. The \emph{Singapore Framework for Dublin Core Application Profiles} \cite{DCMI-Singapore-2008} is a framework for designing metadata and for defining DCAPs. The framework comprises descriptive components that are necessary or useful for documenting DCAPs.

>>>>

\section{Motivation}
\label{sec:motivation}
Even with an upcoming W3C recommendation, it can be expected that several constraint languages will be used in practice in future -- consider the situation in the XML world, where a standardized schema language was available from the beginning and yet additional ways to formulate and check constraints have been created. Therefore, 
semantically equivalent \emph{specific constraints} represented in different languages will exist.
This raises two questions: 
\begin{enumerate}
 \item How can we ensure that two semantically equivalent constraints are actually validated consistently?
 \item How can we support the transformation of semantically equivalent constraints from one contraint language to another?
\end{enumerate}

@@

\subsection{\emph{Simple Constraints}}
In this and the following section, we provide examples for the representation of constraint types within the framework.

@@

There are simple constraints which are not expressible in DL but can still be described using the RDF-CV such as constraints of the type \emph{literal pattern matching} (\emph{R-44}) which restrict literals to match given patterns. The \emph{universal quantification (R-91)} {\small\ms{Book $\sqsubseteq$ $\forall$ identifier.ISBN}} ensures that books can only have valid \emph{ISBN} identifiers, i.e., strings that match a given regular expression.
%Even though, the restriction of the datatype \emph{ISBN} cannot be expressed in DL, \emph{OWL 2 DL} can be used to express the \emph{literal pattern matching} constraint:
%\begin{ex}
%ISBN a RDFS:Datatype ; owl:equivalentClass [ a RDFS:Datatype ;
    %owl:onDatatype xsd:string ; 
    %owl:withRestrictions ([ xsd:pattern "^\d{9}[\d|X]" ])] .
%\end{ex} 
%The first OWL 2 axiom explicitly declares {\em ISBN} to be a datatype. %\er{Why start from second and then first? Also better we say "owl axiom" when we talk axiom of owl, to prevent ambiguity.}\tb{resolved}
%The second OWL 2 axiom defines {\em ISBN} as an abbreviation for a datatype restriction on {\em xsd:string}. 
%The datatype {\em ISBN} can be used just like any other datatype like in the \emph{universal restriction} above.
%There are multiple use cases associated with the requirement to match literals according to given patterns (\emph{Literal Pattern Matching}\footnote{Corresponds to \emph{R-44-PATTERN-MATCHING-ON-RDF-LITERALS}}).
%The enterprise vessel, e.g.,  can only have the registry numbers "NCC-1701", "NCC-1701-A", "NCC-1701-B", "NCC-1701-C", "NCC-1701-D", or "NCC-1701-E".
%The universal restriction can be represented in DL:
%\ms{Enterprise $\sqsubseteq$ $\forall$ registryNumber.RegistryNumber}.
%The restriction of the datatype \ms{RegistryNumber}, however, cannot be expressed in DL, but OWL 2 DL can be used anyway to express the literal pattern matching constraint:
%
%\begin{ex}
%RegistryNumber
    %a rdfs:Datatype ; owl:equivalentClass [ a rdfs:Datatype ;
        %owl:onDatatype xsd:string ;
        %owl:withRestrictions ( [ xsd:pattern "NCC-1701([-][A-E])?" ] ) ] .
%\end{ex}
%The second axiom defines \ms{RegistryNumber} as an abbreviation for a datatype restriction on \ms{xsd:string}. 
%The first axiom explicitly declares \ms{RegistryNumber} to be a datatype. 
%The datatype \ms{RegistryNumber} can be used just like any other datatype like in the universal restriction above.
%The \emph{literal pattern matching} constraint validates \emph{ISBN} literals according to the regular expression causing constraint violations for triples which do not not match. 
%\ms{Janeway commandsEnterprise Voyager} and \ms{Voyager registryNumber "NCC-74656"\textasciicircum{}\textasciicircum{}RegistryNumber}, 
%but not for the triples \ms{Picard commandsEnterprise Enterprise} and \ms{Enterprise registryNumber "NCC-1701-E"\textasciicircum{}\textasciicircum{}RegistryNumber}.

Even though, constraints of the type \emph{literal pattern matching} cannot be expressed in DL, OWL 2 can be used to formulate the constraint:

@@

Table \ref{tab:simple-constraint-not-expressible-in-dl)} presents (1) in the first line how the literal pattern matching simple constraint which is not expressible in DL and (2) in the second line how the universal quantification complex constraint which is expressible in DL are represented using the RDF-CV. Thereby, the context class \emph{ISBN}, whose instances must satisfy the simple constraint, is reused within the list of classes the complex constraint refers to. The literal pattern matching constraint type introduces the constraining element \emph{REGEX} whose validation has to be implemented once like for any other constraining element.
%In general, validation has to be implemented once for each \emph{generic constraint type} which is not expressible in DL. 
%\begin{gcotable}
%property & Enterprise & registryNumber & - & RegistryNumber & $\forall$ & - \\
%class & RegistryNumber & - & - & xsd:string & REGEX & 'NCC-1701([-][A-E])?' \\
%\end{gcotable}
\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Simple Constraints which are not Expressible in DL}
	\label{tab:simple-constraint-not-expressible-in-dl)}
	\centering
		\begin{tabular}{c|c|c|c|c|c}
      \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
      ISBN & - & - & xsd:string & REGEX & '\string^\text{$\backslash$d$\{9\}$[$\backslash$d$\mid$X]}\$' \\
      Book & identifier & - & ISBN & universal quantification & - \\
		\end{tabular}
\end{table} %\tb{Erman, can you please format the regular expression in the table above correctly?}

@@

\subsection{Mapping implementation}

If a language developer would like to use the framework for the implementation of a constraint language, the process is straight-forward. For a language construct, the corresponding constraint type has to be identified. Again we use the \emph{minimum qualified cardinality restriction (R-75)} type which is supported in OWL 2:

@@

From Table~\ref{tab:property-constraint-cardinality-restriction}, we know the representation in our framework, which corresponds to the following 
RDF representation using RDF-CV:

@@

The mapping simply constructs this representation out of the OWL2 representation using a SPARQL CONSTRUCT query:

@@

SPIN is used to execute the mapping, the property \emph{spin:rule} links an rdfs:Class with SPARQL CONSTRUCT queries. Each query defines an inference rule that is applied to all instances of the associated class and its subclasses. The inference rule defines how additional triples can be inferred from what is stated in the WHERE clause. For each binding of the pattern in the WHERE clause of the rule, the triple templates from the CONSTRUCT clause are instantiated and added as inferred triples to the underlying model. At query execution time, the SPARQL variable ?this is bound to the current instance of the class. As each resource per default is assigned to the class owl:Thing, this inference rule is evaluated for each subject of the input RDF graph.

The framework is implemented in exactly the same way by providing another mapping to create the actual SPIN/SPARQL query that validates the constraints and produces constraint violation messages if a constraint is violated, as described in our previous paper about the DSP implementation \cite{BoschEckert2014-2}.\footnote{At the time of this writing, not all mappings for the constraint types are implemented, but of course the implementations can be complemented and adapted to own requirements, as needed. The most recent implementation can be found here: \url{https://github.com/boschthomas/rdf-validation/blob/master/SPIN/RDF-CV-2-SPIN.ttl}.}

\subsection{Constraint transformation}

As stated in Section~\ref{sec:motivation}, we see a huge potential in the possibility to transform semantically equivalent constraints from one high level constraint language to another via the RDF-CV representation, to avoid that every possible combination of constraint languages has to be mapped separately. The following SPIN inference rule exemplifies this approach and provides a mapping from RDF-CV back to an OWL2 constraint:

@@

It can be seen that the mapping is quite similar to the first mapping and basically simply switches the CONSTRUCT and WHERE part of the query, with slight adjustment in the structure of the variables. Potentially an even simpler representation for the mapping could be found that would enable the creation of forward and backward mappings out of it. We didn't investigate this further, though, and it is not yet clear if there can be cases where the backward mapping is more different.

@@

The \emph{Dublin Core Application Profile (DCAP)} and \emph{Bibframe Profiles} are approaches to specify \emph{profiles} for application-specific purposes. The term \emph{profile} is widely used to refer to a document that describes how standards or specifications are deployed to support the requirements of a particular application, function, community, or context. In the metadata community, the term \emph{application profile} has been applied to describe the tailoring of standards for specific applications. A \emph{Dublin Core Application Profile (DCAP)} \cite{DCMI-DCAP-2009} defines metadata records which meet specific application needs while providing semantic interoperability with other applications on the basis of globally defined vocabularies and models. The \emph{Singapore Framework for Dublin Core Application Profiles} \cite{DCMI-Singapore-2008} is a framework for designing metadata and for defining DCAPs. The framework comprises descriptive components that are necessary or useful for documenting DCAPs.
